### 2016-04-27 text/ms.Rmd
Edit main text

<!-- TODO
- [ ] add discussion of uncertainty (CIs) and ML (and incorporating uncertainty into pred.)
- [x] add back in the roc plot and sentence in the main text
- [ ] rerun the main simulation analysis because the equations of change for the slope
- [ ] manually adjusted the labels in figure 4
- [x] update the magic number get inserted to use the equation that only has the basic model
-->

@@

\noindent
Alternative 2: Ensembles of data-limited stock-assessments can improve accuracy
and reduce bias of $B/B_\mathrm{MSY}$ estimates

@@

\noindent \textbf{Abstract:} Fishery managers must often reconcile conflicting
estimates of population status and trend. One solution is to average
predictions using an ensemble method, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used in climate and weather forecasting, may
provide a superior solution. Superensembles use predictions from multiple
models as covariates in an additional model. By training this superensemble
model using known data, we can then use it to predict unknown situations. We
evaluated the potential for ensemble and superensemble models ('ensemble
methods') to improve estimates of population status and trend for fisheries.
We fit four widely applicable data-limited models that estimate stock biomass
relative to the equilibrium biomass at maximum sustainable yield
($B/B_\mathrm{MSY}$). We combined the models' estimates to build four ensemble
methods: an ensemble average and three superensembles (a linear model,
a random forest, and a boosted regression tree). We used the ensemble methods
to predict both (i) fishery status and (ii) recent trends of
$B/B_\mathrm{MSY}$. We trained our models on a simulated dataset of 5760
stocks and tested them with cross-validation and against a global database of
`r ram_stocks_n` real stock assessments. We found that ensemble methods
substantially improved estimates of population status and trend. Random forest
and boosted regression trees performed the best for estimating population
status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order
correlation between predicted and true status improved from `r
mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range` across models?, and
bias (median proportional error) declined from `r mean_sim$ind_mre_range` to
`r mean_sim$mach_mre_range`. We found similar improvements when predicting the
trend of status and when applying the simulation-trained superensembles to
catch data for global fish stocks. Ensemble methods can improve estimates of
status and trends; however, they must be tested, formed from a diverse set of
accurate models, and built on a dataset representative of the populations to
which they are applied. *TODO: Still need to cut ~30 words*

@@

Status and trend are two of the most fundamental values to quantify in the
management of ecological populations [e.g. @iucn2015]. However, managers are
often faced with reconciling multiple uncertain and potentially conflicting
estimates of status and trend [e.g. @brodziak2010; @branch2011; @deroba2015].
For example, one models may suggest a population is at risk and declining in
abundance while others may suggest it is not at risk and stable.

One solution is to take the average or weighted average of several model
predictions, i.e. an ensemble. Such ensembles are typically more accurate and
less biased than individual model estimates and can integrate uncertainty in
model structure, initial conditions, and parameter estimation
[@dietterich2000; @araujo2007]. Ensembles are superior to individual models
for at least three reasons: (1) statistically by averaging across models and
therefore being less likely to pick the "wrong" model, (2) computationally by
reducing the risk of getting stuck in a local optima, and (3)
representationally by expanding the range of hypotheses explored
[@dietterich2000]. This approach forms the basis of many machine learning
methods [e.g. @dietterich2000], has helped reconcile climate forecasts from
dozens of models [e.g. @murphy2004; @tebaldi2007; @ipcc2013], and even
improved early warning signs of malaria outbreaks [@thomson2006]. In ecology,
ensemble methods are sometimes used to improve species distribution modelling
[e.g. @araujo2007; @breiner2015] and indeed have been used to combine
estimates of population status and trend [e.g. @brodziak2010].

@@

Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$
based on the limited data available for the majority of fish stocks: (1)
a time series of the total biomass of catch and (2) a basic understanding of
population productivity [e.g. @vasconcellos2005; @martell2013]. Recently,
@rosenberg2014 investigated the performance of four data-limited models
through a large-scale simulation experiment. Three of these models were based
on Schaefer (logistic) biomass dynamics and one was an empirical model fitted
to more data-rich stock assessment output. The four models frequently
disagreed about population status, no one model had strong performance across
all fish stocks, and some models performed better than others depending on
circumstances (e.g. Fig. \ref{motivate}).

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity
$K$, and $C$ represents catch. The fourth model is an empirically derived model
based on the RAM Legacy Stock Assessment Database [@ricard2012]. @rosenberg2014
provide a full background on these four methods and code to fit all the models
is available in an accompanying package `datalimited` for the statistical
software R <https://github.com/datalimited/datalimited> (*Private while in
review*). In summary:

@@

We first developed and tested ensemble methods on a fully factorial simulated
dataset of fisheries [@rosenberg2014] with known status. These simulated data
included the following scenarios: three fish life histories: small pelagic,
demersal, and large pelagic; three levels of initial biomass depletion
compared to carrying capacity: biomass at 100%, 70%, and 40% of carrying
capacity; and four exploitation settings: (1) a constant exploitation rate,
(2) an exploitation rate coupled with biomass to mimic an open-access
single-species fishery, (3) a scenario where exploitation rate increased
continuously, and (4) a 'roller-coaster' scenario where the exploitation
rate increased and then decreased. Process noise (recruitment variability;
i.e. unexplained variability in population dynamics) was introduced to the
models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$,
and was either uncorrelated through time or had first-order autoregressive
correlation of $0.6$. The simulation also included a scenario with $N(0,
0.2^2)$ measurement error around log(catch) and one scenario without
measurement error. @rosenberg2014 ran ten iterations for each combination
of factors adding stochastic draws of recruitment and catch-recording
variability each time to generate a total of 5760 stocks. The simulation
models were built in the `FLR` packages [@kell2007,
<http://flr-project.org>] for the statistical software R [@r2015]. Code to
generate the simulations is available at
<https://github.com/iagomosqueira/stockStatusFAO>.

@@

@@

@@

In the case of the RAM Legacy Stock Assessment Database, we used superensembles
trained on the entire simulation dataset. However, since mPRM is built on the
same stock assessment database, we applied three-fold cross-validation to the
data underlying the mPRM model so that the dataset with which mPRM was trained
(for the individual model and superensemble) was separate from the dataset with
which it was tested.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g. @walther2005]. Here, we
measure proportional error, defined as $(\hat{\theta} - \theta)/\abs{\theta}$,
where $\hat{\theta}$ and $\theta$ represent estimated and 'true' (or
stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median
proportional error to measure bias, median absolute proportional error to
measure accuracy, and Spearman's rank-order correlation between predicted and
'true' values to measure the ability to correctly rank populations. When
testing with the RAM Legacy Stock Assessment database, we treated the
estimates from these data-rich stock assessments as known without error. Thus,
any error in the stock assessment estimates of the mean or slope of
$B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for
each of the four data-limited models and the ensembles. Code to reproduce our
analysis and an R package to fit the data-limited assessment models are
available at <https://github.com/datalimited/ensembles> (*Private while in
review*).

>>>>

<!-- TODO
- [ ] add discussion of uncertainty (CIs) and ML (and incorporating uncertainty into pred.)
- [x] add back in the roc plot and sentence in the main text
- [x] rerun the main simulation analysis because the equations of change for the slope
- [ ] manually adjusted the labels in figure 4
- [x] update the magic number get inserted to use the equation that only has the basic model
- [ ] check figure order
- [ ] check all remaining in-text todo items
-->

@@

\noindent
Alternative 2: Ensembles of data-limited stock assessments improve accuracy
and reduce bias of $B/B_\mathrm{MSY}$ estimates

@@

\noindent \textbf{Abstract:} Fishery managers must often reconcile conflicting
estimates of population status and trend. One solution is to average
predictions using an ensemble method, but this approach ignores covariance
between models and may not optimally reduce bias and improve accuracy.
Superensemble models, commonly used in climate and weather forecasting, may
provide a superior solution. Superensembles use predictions from multiple
models as covariates in an additional model. By training this superensemble
model using known data, we can then use it to predict unknown situations. We
evaluated the potential for ensemble and superensemble models ('ensemble
methods') to improve estimates of population status and trend for fisheries.
We fit four widely applicable data-limited models that estimate stock biomass
relative to the equilibrium biomass at maximum sustainable yield
($B/B_\mathrm{MSY}$). We combined the models' estimates to build four ensemble
methods: an ensemble average and three superensembles (a linear model,
a random forest, and a boosted regression tree). We used the ensemble methods
to predict both (i) fishery status and (ii) recent trends of
$B/B_\mathrm{MSY}$. We trained our models on a simulated dataset of 5760
stocks and tested them with cross-validation and against a global database of
`r ram_stocks_n` real stock assessments. We found that ensemble methods
substantially improved estimates of population status and trend. Random forest
and boosted regression trees performed the best for estimating population
status: accuracy improved
`r mean_sim$mach_vs_ind_mare_fold`%, rank-order
correlation between predicted and true status improved from
`r mean_sim$ind_corr_range` to
`r mean_sim$mach_corr_range` across models, and
bias (median proportional error) declined from
`r mean_sim$ind_mre_range` to
`r mean_sim$mach_mre_range`. We found similar improvements when predicting the
trend of status and when applying the simulation-trained superensembles to
catch data for global fish stocks. Ensemble methods can improve estimates of
status and trends; however, they must be tested, formed from a diverse set of
accurate models, and built on a dataset representative of the populations to
which they are applied. *TODO: Still need to cut ~30 words*

@@

Status and trend are two of the most fundamental values to quantify in the
management of ecological populations [e.g. @iucn2015]. However, managers are
often faced with reconciling multiple uncertain and potentially conflicting
estimates of status and trend [e.g. @brodziak2010; @branch2011; @deroba2015].
For example, one model may suggest a population is at risk and declining in
abundance while others may suggest it is not at risk and stable.

One solution is to take the average or weighted average of several model
predictions, i.e. an ensemble. Such ensembles are typically more accurate and
less biased than individual model estimates and can integrate uncertainty in
model structure, initial conditions, and parameter estimation
[@dietterich2000; @araujo2007]. Ensembles are superior to individual models
in at least three ways: (1) statistically by averaging across models and
therefore being less likely to pick the "wrong" model, (2) computationally by
reducing the risk of getting stuck in a local optima, and (3)
representationally by expanding the range of hypotheses explored
[@dietterich2000]. This approach forms the basis of many machine learning
methods [e.g. @dietterich2000], has helped reconcile climate forecasts from
dozens of models [e.g. @murphy2004; @tebaldi2007; @ipcc2013], and even
improved early warning signs of malaria outbreaks [@thomson2006]. In ecology,
ensemble methods are sometimes used to improve species distribution modelling
[e.g. @araujo2007; @breiner2015] and indeed have been used to combine
estimates of population status and trend [e.g. @brodziak2010].

@@

Therefore, a number of models have been proposed to assess $B/B_\mathrm{MSY}$
based on the limited data available for the majority of fish stocks: (1) a time
series of the total biomass of catch and (2) a basic understanding of
population productivity [e.g. @vasconcellos2005; @martell2013]. Recently,
@rosenberg2014 investigated the performance of four data-limited models through
a large-scale simulation experiment. Three of these models were based on
Schaefer (logistic) biomass dynamics and one was an empirical model fitted to
more data-rich stock-assessment output. The four models frequently disagreed
about population status (e.g. Fig. \ref{motivate}), no one model had strong
performance across all fish stocks, and some models performed better than
others depending on circumstances.

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity
$K$, and $C$ represents catch. The fourth model is an empirically derived model
based on the RAM Legacy Stock Assessment Database [@ricard2012]. @rosenberg2014
provide a full background on these four methods and code to fit all the models
is available in an accompanying package `datalimited` for the statistical
software R [@r2015] (<https://github.com/datalimited/datalimited>) (*Private while in
review*). In summary:

@@

We first developed and tested ensemble methods on a fully factorial simulated
dataset of fisheries [@rosenberg2014] with known status. These simulated data
included the following scenarios: three fish life histories: small pelagic,
demersal, and large pelagic; three levels of initial biomass depletion
compared to carrying capacity: biomass at 100%, 70%, and 40% of carrying
capacity; and four exploitation settings: (1) a constant exploitation rate,
(2) an exploitation rate coupled with biomass to mimic an open-access
single-species fishery, (3) a scenario where exploitation rate increased
continuously, and (4) a 'roller-coaster' scenario where the exploitation
rate increased and then decreased. Process noise (recruitment variability;
i.e. unexplained variability in population dynamics) was introduced to the
models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$,
and was either uncorrelated through time or had first-order autoregressive
correlation of $0.6$. The simulation also included a scenario with $N(0,
0.2^2)$ measurement error around log(catch) and one scenario without
measurement error. @rosenberg2014 ran ten iterations for each combination
of factors adding stochastic draws of recruitment and catch-recording
variability each time to generate a total of 5760 stocks. The simulation
models were built in the `FLR` packages [@kell2007,
<http://flr-project.org>] for the statistical software R. Code to
generate the simulations is available at
<https://github.com/iagomosqueira/stockStatusFAO>.

@@

TODO: standardize notation of the following with figure 2

@@

TODO add short description of spectral densities

@@

In the case of the RAM Legacy Stock Assessment Database, we used superensembles
trained on the entire simulation dataset. However, since mPRM is built on the
same stock-assessment database, we applied three-fold cross-validation to the
data underlying the mPRM model so that the dataset with which mPRM was trained
(for the individual model and superensemble) was separate from the dataset with
which it was tested.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g. @walther2005]. Here, we
measure proportional error, defined as $(\hat{\theta} - \theta)/\abs{\theta}$,
where $\hat{\theta}$ and $\theta$ represent estimated and 'true' (or
stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median
proportional error to measure bias, median absolute proportional error to
measure accuracy, and Spearman's rank-order correlation between predicted and
'true' values to measure the ability to correctly rank populations. When
testing with the RAM Legacy Stock Assessment database, we treated the
estimates from these data-rich stock assessments as known without error. Thus,
any error in the stock-assessment estimates of the mean or slope of
$B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for
each of the four data-limited models and the ensembles. Code to reproduce our
analysis is available at <https://github.com/datalimited/ensembles> (*Private while in
review*).
